### 2023-08-21 proofs/tex/equiv-lemmas.tex
WIP: description, notation

\begin{lemma}[Declarative equivalence is transitive]
  \hfill
  \label{lemma:decl-equiv-transitivity}
  \begin{itemize}
    \item[$+$] if $[[iP1 ≈ iP2]]$ and $[[iP2 ≈ iP3]]$ then $[[iP1 ≈ iP3]]$,
    \item[$-$] if $[[iN1 ≈ iN2]]$ and $[[iN2 ≈ iN3]]$ then $[[iN1 ≈ iN3]]$.
  \end{itemize}
\end{lemma}
\begin{proof}
  We prove it by
  $\size{[[iP1 ≈ iP2]]} + \size{[[iP2 ≈ iP3]]}$
  and mutually, $\size{[[iN1 ≈ iN2]]} + \size{[[iN2 ≈ iN3]]}$, where
  by size we mean the size of the nodes in the corresponding inference tree.

@@

\begin{lemma}[Algorithmization of declarative equivalence]
  \label{lemma:decl-equiv-algorithmization}
  Declarative equivalence is equality of normal forms. 
  \begin{itemize}
    \item[$+$] $[[iP ≈ iQ]] \iff [[nf(iP) = nf(iQ)]]$,
    \item[$-$] $[[iN ≈ iM]] \iff [[nf(iN) = nf(iM)]]$.
  \end{itemize}
\end{lemma}
\begin{proof} \hfill
  \begin{itemize}
    \item[$+$] Let us prove both directions separately.
    \begin{itemize}
      \item[$\Rightarrow$] 
        exactly by \cref{lemma:normalization-completeness},
      \item[$\Leftarrow$] 
        from \cref{lemma:normalization-soundness}, we know
        $[[iP ≈ nf(iP)]] = [[nf(iQ) ≈ iQ]]$, then by transitivity (\cref{lemma:decl-equiv-transitivity}),
        $[[iP ≈ iQ]]$.
    \end{itemize}
    \item[$-$] The proof is exactly the same.
  \end{itemize}
\end{proof}

@@

\begin{lemma}[Algorithmic type well-formedness is invariant under equivalence]
  \label{lemma:wf-equiv-algo}
  Mutual subtyping implies declarative equivalence.
  \begin{itemize}
  \item[$+$] if $[[uP ≈ uQ]]$ then $[[Γ; Θ ⊢ uP]] \iff [[Γ; Θ ⊢ uQ]]$,
  \item[$-$] if $[[uN ≈ uM]]$ then $[[Γ; Θ ⊢ uN]] \iff [[Γ; Θ ⊢ uM]]$
  \end{itemize}
\end{lemma}
\begin{proof}
  The proof coincides with the proof of \cref{lemma:wf-equiv},
  and adds two cases for equating two positive or two negative metavariables,
  which must be equal by inversion, and thus, 
  $[[Γ; Θ ⊢ α̂±]] \iff [[Γ;  ⊢ α̂±]]$ holds trivially.
\end{proof}

@@

\begin{lemma}[Normalization preserves substitution signature]
  \label{lemma:norm-subst-sig}
  Suppose that $[[σ]]$ is a substitution, $[[Γ1]]$ and $[[Γ2]]$ are contexts.
  Then $[[Γ2 ⊢ σ : Γ1]]$ implies $[[Γ2 ⊢ nf(σ) : Γ1]]$.
\end{lemma}
\being{proof}
  Suppose that $[[α± ∊ {Γ1}]]$. 
  Then by \cref{corollary:wf-nf}, $[[Γ2 ⊢ nf([σ]α±)]] = [[ [nf(σ)]α± ]]$ 
  is equivalent to $[[Γ2 ⊢ [σ]α±]]$.

@@

\begin{corollary}[Normalization preserves well-formedness of algorithmic types]
  \label{corollary:wf-nf-algo}
  \begin{itemize}
    \item[$+$] $[[Γ; Θ ⊢ uP]] \iff [[Γ; Θ ⊢ nf(uP)]]$,
    \item[$-$] $[[Γ; Θ ⊢ uN]] \iff [[Γ; Θ ⊢ nf(uN)]]$
  \end{itemize}
\end{corollary}
\begin{proof}
  Immediately from \cref{lemma:wf-equiv-algo,lemma:normalization-soundness-alg}.
\end{proof}

@@

\begin{corollary}
  \label{corollary:wf-s-nf-algo}
  \hfill \\
  $[[Θ ⊢ uσ]] \iff [[Θ ⊢ nf(σ)]]$
\end{corollary}
\begin{proof}
  The proof is analogous to \cref{corollary:wf-s-nf}.
\end{proof}

@@

>>>>

\begin{lemma}[Declarative equivalence is transitive]
  \hfill
  \label{lemma:decl-equiv-transitivity}
  \begin{itemize}
    \item[$+$] if $[[iP1 ≈ iP2]]$ and $[[iP2 ≈ iP3]]$ then $[[iP1 ≈ iP3]]$,
    \item[$-$] if $[[iN1 ≈ iN2]]$ and $[[iN2 ≈ iN3]]$ then $[[iN1 ≈ iN3]]$.
  \end{itemize}
\end{lemma}
\begin{proof}
  We prove it by
  $\size{[[iP1 ≈ iP2]]} + \size{[[iP2 ≈ iP3]]}$
  and mutually, $\size{[[iN1 ≈ iN2]]} + \size{[[iN2 ≈ iN3]]}$, where
  by size, we mean the size of the nodes in the corresponding inference tree.

@@

\begin{lemma}[Algorithmization of declarative equivalence]
  \label{lemma:decl-equiv-algorithmization}
  Declarative equivalence is the equality of normal forms. 
  \begin{itemize}
    \item[$+$] $[[iP ≈ iQ]] \iff [[nf(iP) = nf(iQ)]]$,
    \item[$-$] $[[iN ≈ iM]] \iff [[nf(iN) = nf(iM)]]$.
  \end{itemize}
\end{lemma}
\begin{proof} \hfill
  \begin{itemize}
    \item[$+$] Let us prove both directions separately.
    \begin{itemize}
      \item[$\Rightarrow$] 
        exactly by \cref{lemma:normalization-completeness},
      \item[$\Leftarrow$] 
        from \cref{lemma:normalization-soundness}, we know
        $[[iP ≈ nf(iP)]] = [[nf(iQ) ≈ iQ]]$, then by transitivity (\cref{lemma:decl-equiv-transitivity}),
        $[[iP ≈ iQ]]$.
    \end{itemize}
    \item[$-$] The proof is exactly the same.
  \end{itemize}
\end{proof}

@@

\begin{lemma}[Algorithmic type well-formedness is invariant under equivalence]
  \label{lemma:wf-equiv-algo}
  Mutual subtyping implies declarative equivalence.
  \begin{itemize}
  \item[$+$] if $[[uP ≈ uQ]]$ then $[[Γ; Ξ ⊢ uP]] \iff [[Γ; Ξ ⊢ uQ]]$,
  \item[$-$] if $[[uN ≈ uM]]$ then $[[Γ; Ξ ⊢ uN]] \iff [[Γ; Ξ ⊢ uM]]$
  \end{itemize}
\end{lemma}
\begin{proof}
  The proof coincides with the proof of \cref{lemma:wf-equiv},
  and adds two cases for equating two positive or two negative metavariables,
  which must be equal by inversion, and thus, 
  $[[Γ; Ξ ⊢ α̂±]] \iff [[Γ; Ξ  ⊢ α̂±]]$ holds trivially.
\end{proof}

@@

\begin{lemma}[Normalization preserves substitution signature]
  \label{lemma:norm-subst-sig}
  Suppose that $[[σ]]$ is a substitution, $[[Γ1]]$ and $[[Γ2]]$ are contexts.
  Then $[[Γ2 ⊢ σ : Γ1]]$ implies $[[Γ2 ⊢ nf(σ) : Γ1]]$.
\end{lemma}
\begin{proof}
  Suppose that $[[α± ∊ {Γ1}]]$. 
  Then by \cref{corollary:wf-nf}, $[[Γ2 ⊢ nf([σ]α±)]] = [[ [nf(σ)]α± ]]$ 
  is equivalent to $[[Γ2 ⊢ [σ]α±]]$.

@@

\begin{corollary}[Normalization preserves well-formedness of algorithmic types]
  \label{corollary:wf-nf-algo}
  \begin{itemize}
    \item[$+$] $[[Γ; Ξ ⊢ uP]] \iff [[Γ; Ξ ⊢ nf(uP)]]$,
    \item[$-$] $[[Γ; Ξ ⊢ uN]] \iff [[Γ; Ξ ⊢ nf(uN)]]$
  \end{itemize}
\end{corollary}
\begin{proof}
  Immediately from \cref{lemma:wf-equiv-algo,lemma:normalization-soundness-alg}.
\end{proof}

@@

\begin{corollary}[Normalization preserves the signature of the algorithmic substitution]
  \label{corollary:norm-subst-sig-algo}
  \hfill \\
  $[[Θ ⊢ uσ]] \iff [[Θ ⊢ nf(uσ)]]$ and
  $[[Γ ⊢ uσ : Ξ]] \iff [[Γ ⊢ nf(uσ) : Ξ]]$.
\end{corollary}
\begin{proof}
  The proof is analogous to \cref{corollary:wf-s-nf}.
\end{proof}

@@

\begin{corollary}[Algorithmic substitution equivalence becomes equality after normalization]
  \label{corollary:subst-subt-equiv-algorithmization}
  Suppose that $[[Θ ⊢ uσ1]]$ and $[[Θ ⊢ uσ2]]$ are algorithmic substitutions.
  Then
  $[[Θ ⊢ uσ1 ≈ uσ2]] \iff [[nf(uσ1)|dom(Θ) = nf(uσ2)|dom(Θ)]]$
  and 
  $[[Γ ⊢ uσ1 ≈ uσ2 : Ξ]] \iff [[nf(uσ1)|Ξ = nf(uσ2)|Ξ]]$.
\end{corollary}

### 2023-08-21 proofs/tex/singularity-lemmas.tex
wip: leftovers
WIP: description, notation

\begin{lemma}[Soundness of Singularity]
    \label{lemma:singularity-soundness}
    Suppose $[[Θ ⊢ SC]]$, and $[[SC singular with uσ]]$. 
    Then $[[Θ ⊢ uσ : SC]]$ and for any 
    $[[uσ']]$ such that $[[Θ ⊢ uσ' : SC]]$,
    $[[Θ ⊢ uσ' ≈ uσ : dom(SC)]]$.
\end{lemma}
\begin{proof}
    Suppose that $[[Θ ⊢ uσ' : SC]]$.
    It means that for every $[[scE]] \in [[SC]]$ restricting $[[α̂±]]$,
    $[[Θ(α̂±) ⊢ [uσ']α̂± : scE]]$ holds.
    $[[SC singular with uσ]]$ means $[[scE singular with [uσ]α̂± ]]$,
    and hence, by \cref{lemma:entry-singularity-completeness},
    $[[ Θ(α̂±) ⊢ [uσ']α̂± ≈ [uσ]α̂±  ]]$ holds.

@@

\begin{lemma}[Completeness of Singularity]
\end{lemma}

\begin{lemma}[Completeness of Singularity]
    \label{lemma:singularity-completeness}
    Suppose there exists $[[Θ ⊢ uσ1]]$ such that
    for any $[[Θ ⊢ uσ]]$, $[[Θ ⊢ uσ : SC]]$ implies $[[Θ ⊢ uσ ≈ uσ1 : varset]]$.
    Then 
    \begin{itemize}
        \item $[[SC|varset singular with uσ0]]$ for some $[[uσ0]]$, and
        \item $[[varset ⊆ dom(SC)]]$.
    \end{itemize} 
\end{lemma}
\begin{proof}
\end{proof}

>>>>

\begin{lemma}[Soundness of Singularity]
    \label{lemma:singularity-soundness}
    Suppose $[[Θ ⊢ SC]]$, and $[[SC singular with uσ]]$. 
    Then $[[(Θ  ⊢  SC) ⊢ uσ]]$ and for any 
    $[[uσ']]$ such that $(Θ  ⊢ SC) ⊢ uσ'$,
    $[[Θ|dom(SC) ⊢ uσ' ≈ uσ]]$.
\end{lemma}
\begin{proof}
    Suppose that $(Θ  ⊢ SC) ⊢ uσ'$.
    It means that for every $[[scE]] \in [[SC]]$ restricting $[[α̂±]]$,
    $[[Θ(α̂±) ⊢ [uσ']α̂± : scE]]$ holds.
    $[[SC singular with uσ]]$ means $[[scE singular with [uσ]α̂± ]]$,
    and hence, by \cref{lemma:entry-singularity-completeness},
    $[[ Θ(α̂±) ⊢ [uσ']α̂± ≈ [uσ]α̂±  ]]$ holds.

@@

\begin{lemma}[Completeness of Singularity]
    \label{lemma:singularity-completeness}
    For a given $[[Θ ⊢ SC]]$
    and $[[varset ⊆ dom(Θ)]]$,
    suppose that all the substitutions satisfying $[[SC]]$ are equivalent,
    in other words, suppose that there exists $[[Θ ⊢ uσ1]]$ such that
    for any $[[Θ ⊢ uσ]]$, $[[(Θ  ⊢  SC) ⊢ uσ]]$ implies $[[Θ|varset ⊢ uσ ≈ uσ1]]$.
    Then 
    \begin{itemize}
        \item $[[SC|varset singular with uσ0]]$ for some $[[uσ0]]$, and
        \item $[[varset ⊆ dom(SC)]]$.
    \end{itemize} 
\end{lemma}
\begin{proof}
    First, let us suppose that $[[varset ⊆ dom(SC)]]$ does not hold.
    It means that there exists $[[ α̂±[Γ] ∊ Θ]]$ and $[[α̂± ∉ dom(SC)]]$

    By \cref{lemma:constraint-sat}, let us take an arbitrary 
    $[[uσ']]$ satisfying $[[SC]]$ (that is $[[(Θ  ⊢  SC) ⊢ uσ']]$).
    Let us extend $[[uσ']]$ to $[[uσ'']]$ so that $[[Θ ⊢ uσ'']]$.

    Although $[[uσ']]$ might not be defined on

    such that for any $[[Θ ⊢ uσ]]$, $[[(Θ  ⊢  SC) ⊢ uσ]]$ implies 
    $[[ Γ ⊢ [uσ]α̂± ≈ [uσ1]α̂± ]]$.
    Let us take

    for any $[[Θ ⊢ uσ]]$, $[[(Θ  ⊢  SC) ⊢ uσ]]$ implies $[[Θ|varset ⊢ uσ ≈ uσ1]]$.

        \item [$-$] Suppose that there exists $[[iN]]$ well-formed in $[[Γ]]$ 
            such that for any $[[iN']]$ well-formed in $[[Γ]]$,
            $[[Γ ⊢ iN' : scE]]$ implies $[[Γ ⊢ iN' ≈ iN]]$. 
            Then $[[scE singular with nf(iN)]]$.
        \item [$+$] Suppose that there exists $[[iP]]$ well-formed in $[[Γ]]$ 
            such that for any $[[iP']]$ well-formed in $[[Γ]]$,
            $[[Γ ⊢ iP' : scE]]$ implies $[[Γ ⊢ iP' ≈ iP]]$. 
            Then $[[scE singular with nf(iP)]]$ .

### 2023-08-21 proofs/tex/subst-lemmas.tex
WIP: description, notation

\begin{lemma}[Substitutions equivalent on the metavariables]
  \label{lemma:subst-equiv-metavar}
  Suppose that $[[Γ ⊢ Θ]]$, $[[uσ1]]$ and $[[uσ2]]$ are substitutions 
  of signature $[[Θ ⊢ uσi]]$.
  Then 
  \begin{enumerate}
    \item [$+$] for a type $[[Γ; Θ ⊢ uP]]$, if $[[Γ ⊢ [uσ1]uP ≈ [uσ2]uP]]$ then
      $[[Θ ⊢ uσ1 ≈ uσ2 : uv uP]]$;
    \item [$-$] for a type $[[Γ; Θ ⊢ uN]]$, if $[[Γ ⊢ [uσ1]uN ≈ [uσ2]uN]]$ then
      $[[Θ ⊢ uσ1 ≈ uσ2 : uv uN]]$.
  \end{enumerate}
\end{lemma}
\begin{proof}
  The proof is a trivial structural induction on 
  $[[Γ; Θ ⊢ uP]]$ and mutually, on $[[Γ; Θ ⊢ uN]]$.
\end{proof}

>>>>

\begin{lemma}[Substitutions equal on the metavariables]
  \label{lemma:subst-eq-metavar}
  Suppose that $[[uσ1]]$ and $[[uσ2]]$ are 
  normalized substitutions of signature $[[Θ ⊢ uσi]]$.
  Then 
  \begin{enumerate}
    \item [$+$] for a normalized type $[[Γ; dom(Θ) ⊢ uP]]$, if $[[ [uσ1]uP = [uσ2]uP ]]$ then
      $[[uσ1|(uv uP)  = uσ2|(uv uP)]]$;
    \item [$-$] for a normalized type $[[Γ; dom(Θ) ⊢ uN]]$, if $[[ [uσ1]uN = [uσ2]uN ]]$ then
      $[[uσ1|(uv uN)  = uσ2|(uv uN)]]$.
  \end{enumerate}
\end{lemma}
\begin{proof}
  The proof is a simple structural induction on 
  $[[Γ; dom(Θ) ⊢ uP]]$ and mutually, on $[[Γ; dom(Θ) ⊢ uN]]$.
  Let us consider the shape of $[[uN]]$ (the cases of $[[uP]]$ are symmetric).
  \begin{caseof}
    \item $[[Γ; dom(Θ) ⊢ α̂⁻]]$. Then $[[ [uσ1]α̂⁻ = [uσ2]α̂⁻ ]]$
      implies $[[uσ1|(uv α̂⁻) = uσ2|(uv α̂⁻)]]$ immediately.
    \item $[[Γ; dom(Θ) ⊢ α⁻]]$. Then $[[uv α̂⁻ = ∅]]$, and
      $[[uσ1|(uv α̂⁻)  = uσ2|(uv α̂⁻)]]$ holds vacuously.
    \item $[[Γ; dom(Θ) ⊢ ∀pas.uN]]$. Then we are proving that
      $[[ [uσ1]∀pas.uN = [uσ2]∀pas.uN ]]$ implies $[[uσ1|(uv ∀pas.uN) = uσ2|(uv ∀pas.uN)]]$.
      By definition of substitution and \ruleref{\ottdruleEOneForallLabel}, 
      $[[ [uσ1]uN = [uσ2]uN ]]$ implies $[[uσ1|uv uN = uσ2|uv uN]]$.
      Since $[[∀pas.uN]]$ is normalized, so is $[[Γ, pas; dom(Θ) ⊢ uN]]$, 
      hence, the induction hypothesis is applicable and implies $[[uσ1|uv uN = uσ2|uv uN]]$,
      as required.
    \item $[[Γ; dom(Θ) ⊢ uP → uN]]$. Then we are proving that
      $[[ [uσ1](uP → uN) = [uσ2](uP → uN) ]]$ implies $[[uσ1|(uv uP → uN) = uσ2|(uv uP → uN)]]$.
      By definition of substitution and congruence of equality, 
      $[[ [uσ1](uP → uN) = [uσ2](uP → uN) ]]$
      means $[[ [uσ1]uP = [uσ2]uP ]]$ and $[[ [uσ1]uN = [uσ2]uN ]]$.
      Notice that $[[uP]]$ and $[[uN]]$ are normalized since $[[uP → uN]]$ is normalized, 
      and well-formed in the same contexts.
      This way, by the induction hypothesis, 
      $[[uσ1|(uv uP) = uσ2|(uv uP)]]$ and $[[uσ1|(uv uN) = uσ2|(uv uN)]]$,
      which since $[[uv (uP → uN) = uv uP ∪ uv uN]]$ implies
      $[[uσ1|(uv uP → uN) = uσ2|(uv uP → uN)]]$.
    \item $[[Γ; dom(Θ) ⊢ ↑uP]]$. The proof is similar to the previous case:
      we apply congruence of substitution, equality, and normalization,
      then the induction hypothesis, and then the fact that $[[uv (↑uP) = uv uP]]$.
  \end{caseof}
\end{proof}

\begin{corollary}[Substitutions equivalent on the metavariables]
  \label{lemma:subst-equiv-metavar}
  Suppose that $[[uσ1]]$ and $[[uσ2]]$ are substitutions 
  of signature $[[Θ ⊢ uσi]]$.
  Then 
  \begin{enumerate}
    \item [$+$] for a type $[[Γ; dom(Θ) ⊢ uP]]$, if $[[Θ ⊢ [uσ1]uP ≈ [uσ2]uP]]$ then
      $[[Θ|uv uP ⊢ uσ1 ≈ uσ2]]$;
    \item [$-$] for a type $[[Γ; dom(Θ) ⊢ uN]]$, if $[[Θ ⊢ [uσ1]uN ≈ [uσ2]uN]]$ then
      $[[Θ|uv uN ⊢ uσ1 ≈ uσ2]]$.
  \end{enumerate}
\end{corollary}
\begin{proof}
  First, let us normalize the types and the substitutions, and show that
  the given equivalences and well-formedness properties are preserved. 
  $[[Γ; dom(Θ) ⊢ uP]]$ implies $[[Γ; dom(Θ) ⊢ nf(uP)]]$ by \cref{corollary:wf-nf-algo}.
  $[[Θ ⊢ [uσ1]uP ≈ [uσ2]uP]]$ implies $[[ nf([uσ1]uP) = nf([uσ2]uP)]]$ by 
  \cref{lemma:subt-equiv-algorithmization}.
  Then $[[ nf([uσ1]uP) = nf([uσ2]uP)]]$ implies 
  $[[ [nf(uσ1)]nf(uP) = [nf(uσ2)]nf(uP)]]$ by \cref{lemma:norm-subst-distr}.
  Notice that by \cref{corollary:norm-subst-sig-algo}
  $[[Θ ⊢ uσi]]$ implies $[[Θ ⊢ nf(uσi)]]$.

  This way, by \cref{lemma:subst-eq-metavar}, 
  $[[Θ ⊢ [uσ1]uP ≈ [uσ2]uP]]$ implies 
  $[[nf(uσ1)|(uv nf(uP)) = nf(uσ2)|(uv nf(uP))]]$.
  Then by \cref{lemma:uv-nf}, 
  $[[nf(uσ1)|(uv uP) = nf(uσ2)|(uv uP)]]$,
  and by ${{corollary:subst-subt-equiv-algorithmization}}$,
  $[[Θ|uv uP ⊢ uσ1 ≈ uσ2]]$.

  Symmetrically, 
  $[[Θ ⊢ [uσ1]uN ≈ [uσ2]uN]]$ implies 
  $[[Θ|uv uN ⊢ uσ1 ≈ uσ2]]$.
\end{proof}

### 2023-08-21 proofs/tex/unif-constraint-merge-lemmas.tex
WIP: description, notation

\begin{corollary}
    \label{corollary:unif-merge-soundness}
    Suppose that $[[Θ ⊢ UC1]]$ and $[[Θ ⊢ UC2]]$ 
    are normalized unification constraints.
    If $[[Θ ⊢ UC1 & UC2 = UC]]$ is defined then
    \begin{enumerate}
        \item $[[Θ ⊢ UC]]$ is normalized unification constraint,
        \item for any substitution $[[Θ ⊢ uσ]]$, $[[Θ ⊢ uσ : lift UC]]$ implies 
        $[[Θ ⊢ uσ : lift UC1]]$ and $[[Θ ⊢ uσ : lift UC2]]$.
    \end{enumerate}
\end{corollary}
\begin{proof}
    It is clear that since $[[UC = UC1 ∪ UC2]]$ (by \cref{lemma:unif-merge-soundness}),
    and being normalized means that all entries are normalized,
    $[[UC]]$ is a normalized unification constraint.
    Analogously, $[[Θ ⊢ UC]] = [[UC1 ∪ UC2]]$ holds immediately, 
    since $[[Θ ⊢ UC1]]$ and $[[Θ ⊢ UC2]]$.

    Let us take an arbitrary substitution $[[Θ ⊢ uσ]]$ and assume that 
    $[[Θ ⊢ uσ : lift UC]]$.
    Then $[[Θ ⊢ uσ : lift UCi]]$ holds by definition:
    If $[[ucE]] \in [[lift UCi]] \subseteq [[lift UC1 ∪ lift UC2]] = [[lift UC]]$.
    So $[[Θ(α̂±) ⊢ [uσ]α̂± : ucE]]$ holds.
\end{proof}

@@

\begin{lemma} [Completeness of Unification Constraint Merge] 
    \label{lemma:unif-merge-completeness}
    Suppose that $[[Θ ⊢ UC1]]$ and $[[Θ ⊢ UC2]]$.
    Then for any substitution $[[Θ ⊢ uσ]]$ such that $[[Θ ⊢ uσ : lift UC1]]$ and $[[Θ ⊢ uσ : lift UC2]]$, 
    \begin{enumerate}
        \item $[[Θ ⊢ UC1 & UC2 = UC]]$ is defined and
        \item $[[Θ ⊢ uσ : UC]]$.
    \end{enumerate}
\end{lemma}
\begin{proof}
    The proof repeats the proof of \cref{lemma:merge-completeness} but 
    uses \cref{lemma:unif-entry-merge-completeness} instead of \cref{lemma:entry-merge-completeness}.
\end{proof}

>>>>

\begin{corollary}
    \label{corollary:unif-merge-soundness}
    Suppose that $[[Θ ⊢ UC1]]$ and $[[Θ ⊢ UC2]]$ 
    are normalized unification constraints.
    If $[[Θ ⊢ UC1 & UC2 = UC]]$ is defined then
    \begin{enumerate}
        \item $[[Θ ⊢ UC]]$ is normalized unification constraint,
        \item for any substitution $[[Θ ⊢ uσ]]$, $[[(Θ  ⊢  lift UC) ⊢ uσ]]$ implies 
        $[[(Θ  ⊢  lift UC1) ⊢ uσ]]$ and $[[(Θ  ⊢  lift UC2) ⊢ uσ]]$.
    \end{enumerate}
\end{corollary}
\begin{proof}
    It is clear that since $[[UC = UC1 ∪ UC2]]$ (by \cref{lemma:unif-merge-soundness}),
    and being normalized means that all entries are normalized,
    $[[UC]]$ is a normalized unification constraint.
    Analogously, $[[Θ ⊢ UC]] = [[UC1 ∪ UC2]]$ holds immediately, 
    since $[[Θ ⊢ UC1]]$ and $[[Θ ⊢ UC2]]$.

    Let us take an arbitrary substitution $[[Θ ⊢ uσ]]$ and assume that 
    $[[(Θ  ⊢  lift UC) ⊢ uσ]]$.
    Then $[[(Θ  ⊢  lift UCi) ⊢ uσ]]$ holds by definition:
    If $[[ucE]] \in [[lift UCi]] \subseteq [[lift UC1 ∪ lift UC2]] = [[lift UC]]$.
    So $[[Θ(α̂±) ⊢ [uσ]α̂± : ucE]]$ holds.
\end{proof}

@@

\begin{lemma} [Completeness of Unification Constraint Merge] 
    \label{lemma:unif-merge-completeness}
    Suppose that $[[Θ ⊢ UC1]]$ and $[[Θ ⊢ UC2]]$.
    Then for any substitution $[[Θ ⊢ uσ]]$ such that $[[(Θ  ⊢  lift UC1) ⊢ uσ]]$ and $[[(Θ  ⊢  lift UC2) ⊢ uσ]]$, 
    \begin{enumerate}
        \item $[[Θ ⊢ UC1 & UC2 = UC]]$ is defined and
        \item $[[(Θ  ⊢  UC) ⊢ uσ]]$.
    \end{enumerate}
\end{lemma}
\begin{proof}
    The proof repeats the proof of \cref{lemma:merge-completeness} but 
    uses \cref{lemma:unif-entry-merge-completeness} instead of \cref{lemma:entry-merge-completeness}.
\end{proof}
